---
title: "Life Expectancy - Regression Analysis"
author: "Yevonnael Andrew"
date: "2/16/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

```{r, message=FALSE, warning=FALSE}
library(tidyverse)
library(corrplot)
library(leaps)
library(ggpubr)
library(GGally)
library(DescTools)
```

```{r}
le_data <- read_csv("https://raw.githubusercontent.com/JNYH/Project-Luther/master/df3.csv")
```

```{r}
head(le_data)
```

```{r}
summary(le_data)
```

```{r}
sapply(le_data, function(y) sum(length(which(is.na(y)))))
```

```{r}
cor_mat <- cor(le_data[,-1])
head(cor_mat[1:5, 1:5])
```

```{r, fig.width=12}
corrplot(cor_mat, method="pie", type="lower", addCoef.col = "black")
```

```{r}
le_data_nc <- le_data[,-1]
```

```{r, fig.width=12, warning=FALSE, message=FALSE}
ggpairs(le_data_nc)
```

Model with the most significant variable
```{r}
model_single <- lm(`Life Expectancy` ~ `Birth Rate`, data = le_data_nc)
summary(model_single) # R-squared:  0.5995
```

Model with all variable included
```{r}
model_full <- lm(`Life Expectancy` ~ ., data = le_data_nc)
summary(model_full) # Adjusted R-squared:  0.7369
```

## Feature Selection

1. Stepwise Backward
```{r}
model_bw <- step(model_full, direction = "backward")
summary(model_bw) # Adjusted R-squared:  0.7404
```

2. Stepwise Forward
```{r}
model_none <- lm(`Life Expectancy` ~ 1, data = le_data_nc)
model_fw <- step(model_none, scope = list(lower = model_none, upper = model_full), direction = "forward")
summary(model_fw) # Adjusted R-squared:  0.7404
```

3. Stepwise Both
```{r}
model_bo <- step(model_none, scope = list(lower = model_none, upper = model_full), direction = "both")
summary(model_bo) # Adjusted R-Squared = 0.7404
```

## Using regsubset

```{r}
model_rss <- regsubsets(`Life Expectancy` ~ ., data = le_data_nc, nbest = 2)
plot(model_rss, scale = "adjr2")
```

```{r}
sum_model_rss <- summary(model_rss)
names(sum_model_rss)
```

```{r}
which.max(sum_model_rss$adjr2)
```

```{r}
which.min(sum_model_rss$bic)
```

```{r}
sum_model_rss$which[9,]
```

```{r}
sum_model_rss$which[5,]
```

```{r}
lm_model_rss_r2 <- lm(`Life Expectancy` ~ `Birth Rate` + EPI + `Health Expenditure` + `Pop Density` + `Stroke Rate`, data = le_data)
summary(lm_model_rss_r2) # Adjusted R-squared:  0.7406
```

```{r}
lm_model_rss_bic <- lm(`Life Expectancy` ~ `Birth Rate` + EPI + `Stroke Rate`, data = le_data)
summary(lm_model_rss_bic) # Adjusted R-squared:  0.739
```


```{r}
plot1 <- ggplot(model_bo, aes(.resid)) + geom_histogram(bins = 50)
plot2 <- ggplot(lm_model_rss_r2, aes(.resid)) + geom_histogram(bins = 50)
plot3 <- ggplot(lm_model_rss_bic, aes(.resid)) + geom_histogram(bins = 50)
plot4 <- ggplot(model_full, aes(.resid)) + geom_histogram(bins = 50)

ggarrange(plot1, plot2, plot3, plot4)
```

```{r}
shapiro.test(model_bo$residuals)
```

## Creating a New Variable

```{r, warning=FALSE, message=FALSE}
country_names <- read.csv("continent_country.csv", sep = ";", stringsAsFactors = FALSE)
colnames(country_names) <- c("Continent_Name", "Country_Name")
african_country <- country_names %>% filter(Continent_Name == "Africa")
asian_country <- country_names %>% filter(Continent_Name == "Asia")
europe_country <- country_names %>% filter(Continent_Name == "Europe")
antarctica_country <- country_names %>% filter(Continent_Name == "Antarctica")
oceania_country <- country_names %>% filter(Continent_Name == "Oceania")
northamerica_country <- country_names %>% filter(Continent_Name == "North America")
southamerica_country <- country_names %>% filter(Continent_Name == "South America")
```

```{r}
african_country <- tolower(unlist(strsplit(african_country$Country_Name, ",")))
asian_country <- tolower(unlist(strsplit(asian_country$Country_Name, ",")))
europe_country <- tolower(unlist(strsplit(europe_country$Country_Name, ",")))
antarctica_country <- tolower(unlist(strsplit(antarctica_country$Country_Name, ",")))
oceania_country <- tolower(unlist(strsplit(oceania_country$Country_Name, ",")))
northamerica_country <- tolower(unlist(strsplit(northamerica_country$Country_Name, ",")))
southamerica_country <- tolower(unlist(strsplit(southamerica_country$Country_Name, ",")))
```

```{r}
le_data <- le_data %>%
  mutate(Continent = ifelse(Country %in% african_country, "Africa",
                     ifelse(Country %in% asian_country, "Asia", 
                     ifelse(Country %in% europe_country, "Europe",
                     ifelse(Country %in% northamerica_country, "North America",
                     ifelse(Country %in% southamerica_country, "South America", 
                     ifelse(Country %in% oceania_country, "Oceania",
                     ifelse(Country == "guernsey (channel islands, uk)", "Europe",
                     ifelse(Country == "jersey (channel islands, uk)", "Europe",
                     ifelse(Country == "saint helena, ascension and tristan da cunha (uk)", "Africa",
                     "NONE"))))))))))
```

I am sure that Continent affect the life expectancy. Let's create a boxplot to ensure it.

```{r}
le_data %>%
  ggplot(aes(x=Continent, y=`Life Expectancy`)) + geom_boxplot()
```

```{r}
model_full_continent <- lm(`Life Expectancy` ~ ., data = le_data[,-1])
summary(model_full_continent)
```

## Detecting outliers

```{r, fig.width=10}
par(mfrow = c(2,2))
plot(model_full_continent)
```


```{r}

```

## Winsorizing Data
We will replace extreme values by less extreme ones.
https://en.wikipedia.org/wiki/Winsorizing
https://www.rdocumentation.org/packages/DescTools/versions/0.99.32/topics/Winsorize

```{r}
data_winsorize <- le_data
data_winsorize$`Dengue Cases` <- Winsorize(data_winsorize$`Dengue Cases`, probs = c(0.05, 0.95))
data_winsorize$GDP <- Winsorize(data_winsorize$GDP, probs = c(0.05, 0.95))
data_winsorize$`Pop Density` <- Winsorize(data_winsorize$`Pop Density`, probs = c(0.05, 0.95))
data_winsorize$Population <- Winsorize(data_winsorize$Population, probs = c(0.05, 0.95))
data_winsorize$Area <- Winsorize(data_winsorize$Area, probs = c(0.05, 0.95))
```

```{r, fig.width=12, warning=FALSE, message=FALSE}
ggpairs(data_winsorize[,-1])
```

```{r}
model_full <- lm(`Life Expectancy` ~ . ,data = data_winsorize[-64,-1])
shapiro.test(model_full$residuals)
```


## Applying LOG transformation to all features

```{r}
log_data <- sqrt(data_winsorize[-c(66, 51, 75, 214, 158, 120),-c(1, 12, 14)])
log_data <- cbind(log_data, data_winsorize[-c(66, 51, 75, 214, 158, 120),c(12, 14)])
```

```{r, fig.width=12, warning=FALSE, message=FALSE}
ggpairs(log_data[,-1])
```

```{r, fig.width=12}
cor_mat <- cor(log_data[, -c(11, 13)])
corrplot(cor_mat, method="pie", type="lower", addCoef.col = "black")
```

Model with the most significant variable
```{r}
model_single_log <- lm(`Life Expectancy` ~ `Birth Rate`, data = log_data)
summary(model_single_log) # R-squared:  0.5542
```

Model with all variable included
```{r}
model_full_log <- lm(`Life Expectancy` ~ ., data = log_data)
summary(model_full_log) # Adjusted R-squared:  0.7279
```

## Feature Selection

1. Stepwise Backward
```{r}
model_bw_log <- step(model_full_log, direction = "backward")
summary(model_bw_log) # Adjusted R-squared:  0.7311
```

2. Stepwise Forward
```{r}
model_none_log <- lm(`Life Expectancy` ~ 1, data = log_data)
model_fw_log <- step(model_none_log, scope = list(lower = model_none_log, upper = model_full_log), direction = "forward")
summary(model_fw_log) # Adjusted R-squared:  0.7311
```

3. Stepwise Both
```{r}
model_bo_log <- step(model_none_log, scope = list(lower = model_none_log, upper = model_full_log), direction = "both")
summary(model_bo_log) # Adjusted R-Squared = 0.7311
```

```{r}
shapiro.test(model_bo_log$residuals)
```

```{r}
hist(model_bo_log$residuals, breaks = 100)
```

```{r}
log_data <- cbind(log_data, data_winsorize[-c(66, 51, 75, 214, 158, 120),1])

```

```{r}
plot(model_bo_log)
```

```{r}
hist(le_data$`Life Expectancy`)
```


```{r}
car::vif(model_bo_log)
```

```{r}
bptest(model_bo_log)
```

